---
title: "Data Tidying"
author: "CForster"
date: "May 21, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, Libraries}
library(dplyr)
library(tidyr)
```

Read in data from the [KNB]https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1
```{r}
catch_df <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),
                    stringsAsFactors = FALSE)
```
Get fid of summary column and notes column 
Select only keeps the variables you mention
Can also negatively select and say "everything but this"

Consolidating entire pipe 
```{r}
unique(catch_df$notesRegCode)
catch_df <- catch_df %>%  
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum) %>% 
  gather(key = Species, value = Catch, -Region, -Year)
  rename(catch_thousands = Catch)

#catch_df <- catch_df %>%  
#  select(-All, -notesRegCode)
```

Gather and Spread
spread is the opposite of gather and moves your data back to wide format
```{r}
catch_df <- catch_df %>% 
  gather(key = Species, value = Catch, -Region, -Year)

head(catch_df)

# We don't actually want data in wide format
#catch_df %>% 
#  spread(key = Species, value = Catch)
```

Rename columns
new name equals old name 
```{r}
catch_df <- catch_df %>% 
  rename(catch_thousands = Catch)

head(catch_df)
```

Mutate function
new thing equals old thing

diagnosing non-integer values 

Found a typo in original data, corrected it, converted `catch_thousands` column to thousands of fish 
```{r}
test <- as.integer(catch_df$catch_thousands)
which(is.na(test == TRUE))
catch_df[401,]

catch_df <- catch_df %>% 
  mutate(catch_thousands = ifelse(catch_thousands == "I", 1, catch_thousands)) %>% 
  mutate(catch_thousands = as.integer(catch_thousands)) %>% 
  mutate(catch = catch_thousands * 1000)

head(catch_df)
```

# Analysis 

Group_by almost always used with summarize or mutate
Can pass multiple variables to group by 
n() counts number of observations within a group
```{r}
catch_df %>% 
  group_by(Species, Year) %>% 
  summarize(mean_catch = mean(catch), num_obs = n())
```

Filter  
```{r}
catch_df %>% 
  filter(Species == "Chinook") %>% 
  group_by(Region) %>% 
  summarise(mean_catch = mean(catch)) %>% 
  arrange(desc(mean_catch))
```

